---
title: "Burstiness in LDP"
author: " Maddie Meyers and Dan Yurovsky"
date: "`r Sys.Date()`"
output: 
  html_document:
  toc: false
number_sections: false
theme: lumen
toc_float: false
code_folding: hide
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(langcog)
library(stringr) 
library(lme4)
library(directlabels)
library(DT)
library(broom)
library(data.table)
library(ggplot2)
library(forcats)
library(feather)
library(tidytext)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = TRUE, tidy = FALSE)

theme_dy <- function(base_size = 14) 
{
  theme_bw() +   
    ggplot2::`%+replace%`(ggplot2::theme_bw(base_size = base_size),
                          ggplot2::theme(panel.grid = ggplot2::element_blank(), 
                                         legend.position = "none"))
}

theme_set(theme_dy())
```

## R Markdown

Setting up data
```{r load_data}
setwd("~/Documents/Github/gesture")
# Original data
loaded_subjs <- read_feather("feathers/loaded_subjs_full.feather")
# Referents and their frequencies
referents <- read_feather("feathers/referents.feather")
# End result of splitting referents and coding modality
coded_responses <- read_feather("feathers/coded_responses.feather")
```

##Zipfian Distributions 
Tried to create these to show trends in burstiness of the data, but it is hard to gleam any general knowledge from graphs.
```{r zipfs}
person_freqs <- coded_responses %>%
  group_by(subj, age, person, modality, referent) %>%
  summarise(person_modality_n = n()) %>%
  group_by(subj, age, referent) %>%
  mutate(total_n = sum(person_modality_n)) 

ranks <- person_freqs %>%
  arrange(subj, age, desc(total_n)) %>%
  distinct(subj, age, referent, .keep_all = TRUE) %>% 
  group_by(subj, age) %>% 
  mutate(rank = 1:n()) %>%
  select(subj, age, referent, rank)

freq_ranks <- left_join(person_freqs, ranks)

#modality and age
ggplot(filter(freq_ranks, modality != "both"), 
       aes(x = rank, y = person_modality_n, 
                  color = as.factor(age), linetype = modality)) + 
  facet_grid(~person) + 
  #geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() + 
  scale_y_log10() +
  theme_classic()
   
person_zipfs <- freq_ranks %>%  
  filter(modality != "both") %>%
  group_by(person, modality) %>%
  do(tidy(lmer(log(person_modality_n) ~ log(rank)*age + (log(rank) + age|subj), data = .))) 


ggplot(filter(person_zipfs, term %in% c("(Intercept)", "age", "log(rank)", "log(rank):age")),
       aes(x = modality, y = estimate, color = person)) + 
  facet_wrap(~ term, scales = "free") + 
  geom_hline(yintercept = 0, lty = 2) + 
  geom_pointrange(aes(ymin = estimate-std.error, ymax = estimate+std.error),
                      position = position_dodge(.5)) + 
  theme_bw()

person_zipfs <- freq_ranks %>%  
  filter(modality != "both") %>%
  group_by(person, modality) %>%
  do(tidy(lmer(log(person_modality_n) ~ log(rank)*age + (log(rank) + age|subj), data = .))) 


ggplot(filter(person_zipfs, term %in% c("(Intercept)", "age", "log(rank)", "log(rank):age")),
       aes(x = modality, y = estimate, color = person)) + 
  facet_wrap(~ term, scales = "free") + 
  geom_hline(yintercept = 0, lty = 2) + 
  geom_pointrange(aes(ymin = estimate-std.error, ymax = estimate+std.error),
                      position = position_dodge(.5)) + 
  theme_bw()
```

##Main Topic Analysis 
Here, we wanted to see if we could break down the data for if it introduced a new topic or not. The topics were marked by whether the difference between successive mentions of the same topic was greater or less than predicted under a random-chance frequency-dependent model. In this analysis, we underestimated what was included in a topic by marking referents that were mentioned significantly sooner than expected as being part of the same topic. This underestimation was so great that for some referents, even if the difference was 1 it was not classified as in the same topic. 

We then fit a model to see if factors like age, modality, and person (parent/child) were significant in predicting whether a referent was part of the same topic. 

```{r intervals}
intervals <- coded_responses %>%
  group_by(subj, age, referent) %>%
  arrange(time) %>%
  summarise(freq = n(), times = paste0(diff(as.numeric(time)), collapse = " ")) %>%
  unnest_tokens(diff, times)

annotated_intervals <- coded_responses %>%
  group_by(subj, age, referent) %>%
  mutate(freq = n()) %>%
  filter(freq > 1) %>%
  arrange(subj, age, referent, time) %>%
  slice(2:n()) %>%
  ungroup() %>%
  select(time, person, modality, freq_cut) %>%
  bind_cols(intervals, .) %>%
  mutate(diff = as.numeric(diff)) %>%
  group_by(subj, age) %>%
  mutate(total_refs = n(),
         predicted = total_refs/freq,
         prob_extreme = pexp(diff, rate = freq/total_refs, 
                             lower.tail = T, log.p = T),
         old_topic = prob_extreme < log(.05),
         topic = ifelse(prob_extreme < log(.05), "old_topic", NA)) %>% #underestimates new_topic
  ungroup() %>%
  mutate(modality = factor(modality, levels = c("speech", "gesture", "both")))

hist(exp(annotated_intervals$prob_extreme))

old_topic_prob <- annotated_intervals %>%
  group_by(age, person, subj) %>%
  summarise(old_topic = mean(old_topic)) %>%
  summarise(sem = sd(old_topic)/sqrt(n()-1), old_topic = mean(old_topic))
```

```{r models}
old_topic_model <- glmer(old_topic ~ age*person*modality + log(freq)  +
                            (1|subj), 
      family = "binomial", data = annotated_intervals)

old_topic_model_simple <- glmer(old_topic ~ scale(log(age)) * person * modality * scale(log(freq))  +
                            (1|subj), 
      family = "binomial", data = filter(annotated_intervals, modality != "both"))

summary(old_topic_model_simple)

old_topic_model_parent <- glmer(old_topic ~ scale(log(age))  * scale(log(freq)) +
                            (1|subj), 
      family = "binomial", data = filter(annotated_intervals, modality != "both", person == "parent"))

summary(old_topic_model_parent)
```

It is really difficult to represent the trends seen from the above model. Here are a few attempts:

```{r graphs}
ggplot(old_topic_prob, aes(x = age, y = old_topic, color = person, label= person)) + 
  geom_pointrange(aes(ymin = old_topic - sem, ymax = old_topic + sem), position = position_dodge(.25)) + 
  geom_dl(method = "smart.grid")

old_topic_prob <- annotated_intervals %>%
  filter(modality != "both") %>% 
  group_by(age, modality,  person, freq_cut, subj, referent) %>% 
  summarise(old_topic = mean(old_topic)) #%>%
   summarise(old_topic = mean(old_topic))# %>%
  # summarise(old_topic = mean(old_topic)) %>%
  # summarise(old_topic = mean(old_topic))

ggplot(old_topic_prob, aes(x = freq_cut, y = old_topic, color =modality, label = modality,
                         group = modality)) + 
  facet_grid(age~ person) +
  geom_smooth() +
  geom_dl(method = "last.qp")

topic_prop<-annotated_intervals %>%
  mutate(topic=ifelse(prob_extreme > log(.95), "new_topic",ifelse(prob_extreme < log(.05), "old_topic", NA))) %>%
  group_by(age, person, topic) %>%
  summarise(n=n()) %>%
  mutate(prop=n/sum(n))

#Topic (old/new/neither) probability
ggplot(topic_prop, aes(x=age, y=prop, fill=topic))+
  geom_bar(stat="identity",position="dodge")+
  facet_wrap(~person)+
  scale_x_continuous(breaks= c(14, 18, 22, 26, 30, 34))+
  theme(legend.position= "bottom")

#also remember that we only have n=9 @ 30mos
topic<- annotated_intervals %>%
  mutate(topic=ifelse(prob_extreme > log(.95), "new_topic",ifelse(prob_extreme < log(.05), "old_topic", NA))) 

#Topic (old.new.neither), frequency
ggplot(topic, aes(x = age, fill=topic)) + 
  geom_bar(position="dodge") +
  facet_wrap(~person)+
  scale_x_continuous(breaks= c(14, 18, 22, 26, 30, 34)) +
  theme(legend.position= "bottom")
```

```{r- two-way interaction representations}
#age/person/old_topic
person_old_topic<- annotated_intervals%>%
  filter(modality != "both") %>%
  group_by(age, person, old_topic) %>%
  summarize(n=n()) %>%
  mutate(prob=n/sum(n))

ggplot(person_old_topic, aes(x = age, y = prob, fill=person)) + 
  geom_bar(stat="identity", position="dodge") +
  facet_wrap(~old_topic)+
  scale_x_continuous(breaks= c(14, 18, 22, 26, 30, 34)) +
  theme(legend.position= "bottom") + 
  ylab("old_topic probability")

#age/modality/old_topic
modality_old_topic<- annotated_intervals%>%
  filter(modality != "both") %>%
  group_by(age, modality, old_topic) %>%
  summarize(n=n()) %>%
  mutate(prob=n/sum(n)) 

ggplot(modality_old_topic, aes(x = age, y = prob, fill=modality)) + 
  geom_bar(stat="identity", position="dodge") +
  facet_wrap(~old_topic)+
  scale_x_continuous(breaks= c(14, 18, 22, 26, 30, 34)) +
  theme(legend.position= "bottom") + 
  ylab("old_topic probability")

#person/modality (AS EXPECTED)
#modality:gesture reduces the effect of person:parent on prob old_topic to where it is only slightly negative
modality_person_old_topic<- annotated_intervals%>%
  filter(modality != "both") %>%
  group_by(subj, person, modality, old_topic) %>%
  summarize(n=n()) %>%
  mutate(prob=n/sum(n)) 

ggplot(modality_person_old_topic, aes(x = person, y = prob, fill=modality)) + 
  geom_bar(stat="identity", position="dodge") +
  facet_wrap(~old_topic)+
  theme(legend.position= "bottom") + 
  ylab("old_topic probability")
```

```{r- three way interaction representation}
# is this possible? 
#age/modality/person/old_topic
#old_topic=TRUE, see proportion
#age is x axis 

three_way<-annotated_intervals %>%
  filter(modality != "both") %>%
  group_by(age, person, modality, old_topic) %>%
  summarize(n=n()) %>%
  mutate(prob=n/sum(n)) %>%
  filter(old_topic==TRUE)

#does this actually have any meaning? probably not
ggplot(three_way, aes(x=age, y=modality, fill=prob)) +
  geom_tile()+
  scale_x_continuous(breaks= c(14, 18, 22, 26, 30, 34)) +
  facet_grid(~person)+
  theme(legend.position="bottom")

```

Old code that calculated predicted difference by using means. Has been replaced by above interval code
```{r}
intervals %>%
  group_by(freq, age, subj, referent) %>%
  summarise(diff = median(diff-predicted)) %>%
  summarise(diff = mean(diff)) %>%
  summarise(diff = mean(diff)) %>%
  ggplot(aes(x = freq, y = diff, color = as.factor(age))) + 
    geom_point() + 
    geom_smooth(se = FALSE) + 
    scale_x_log10() + 
  geom_hline(yintercept = 0, lty = 2) +
  # scale_y_log10() +
    theme_bw()


diff_lm <- lmer((diff-predicted) ~ log(freq) + age + log(freq)*age +
                   (log(freq)|subj), data = intervals)


intervals %>%
  group_by(freq, age, subj, referent) %>%
  summarise(new_topic = mean(new_topic)) %>%
  summarise(new_topic = mean(new_topic)) %>%
  summarise(new_topic = mean(new_topic)) %>%
  ggplot(aes(x = freq, y = new_topic, color = as.factor(age))) + 
    geom_point() + 
    geom_smooth(se = FALSE, method = "lm") + 
    scale_x_log10() + 
  geom_hline(yintercept = 0, lty = 2) +
  # scale_y_log10() +
    theme_bw()

diff_logit <- glmer(new_topic ~ log(freq) + age + (1|subj),
                   family = "binomial", data = intervals)
```

##Gleitman Plots & Similarity Judgments 
A way of representing how referents shift during a conversation.

It is surprising that so many of the referents fit into these neat packs 
```{r similarity word packs}
#made for subjects 33, 103 
food_words<- c("pepper","strawberry","knife","egg", "pot", "lid","refrigerator", "eggplant", "kiwi", "fork","cup","apple","casserole","lemon","green_bean", "mustard","onion","carrot","soup", "food", "juice", "spoon", "plate", "orange", "popsicle", "cookie", "ice_cream", "milk", "roast", "meat", "plum", "yogurt", "hash_brown", "grape", "hamburger", "pickle", "candy", "soda", "cheese", "turkey", "chocolate", "cheeto", "cereal", "chip", "milk", "shortcake", "jellybean", "coffee", "tea", "sugar", "cheerio", "spatula", "frying_pan", "fry", "milk", "toast", "blueberry", "banana", "potpourri", "ketchup", "cookie_cutter", "oven_mitt", "fridge", "fruit", "bacon", "waffle", "tomato", "teapot", "toaster", "ice_pop", "vanilla", "watermelon", "fruit_snack")

animal_words<-c("pig", "cow", "fish", "horse", "turtle", "duck", "farm", "mouse", "teddy_bear", "peacock", "monkey", "dog", "cat", "bear", "ladybug", "alligator", "toucan", "rabbit", "bug", "cow", "octopus", "wolf", "bird", "giraffe", "tail", "gorilla", "orangutan", "flamingo", "dinosaur", "lion", "elephant", "animal", "walrus", "penguin", "frog", "spider", "bee", "unicorn", "weasel", "zebra", "bunny",  "tiger", "camel", "squirrel", "rooster")

baby_words<-c("baby", "cradle", "stroller", "bottle", "rattle", "diaper", "bib", "pacifier")

body_words<-c("eye", "hand", "foot", "arm", "back", "face", "mouth", "tooth", "leg", "body", "nose", "neck", "knee", "head", "finger", "stomach", "hair", "hip", "cheek", "thumb", "belly_button", "toe", "butt", "shoulder")

#probably abstract (low burstiness) words? in this basic category below. From graphs it seems like not!
basic_words<-c("circle", "ground",  "square", "triangle", "button", "book", "floor", "hole",  "box", "shape", "top", "bottom")

home_words<-c("tv", "couch", "door", "bed", "garbage", "phone", "kitchen", "remote", "vacuum", "towel", "bathroom", "room", "bedroom", "wall", "mop", "potty", "toothpaste", "bathtub", "table", "fireplace", "shelf", "key", "sink", "basement", "dining_room", "blanket", "window", "cupboard", "coathanger", "toilet", "mousepad", "tv", "radio", "tissue", "trash", "light", "paper_towel", "house", "clothesline", "receipt", "roof", "pillow", "chimney")

toy_words<-c("truck", "doll", "jumprope", "toy", "toybox", "puzzle", "puzzle_piece", "draw_board", "slide", "stencil", "block", "crayon", "bubble", "paper", "card", "marker", "train", "rubber_duck", "hammer", "ball", "railroad_track", "jackinthebox", "tower", "castle", "pooh", "water_gun", "pinata", "balloon", "bubble_wand", "bucket", "playpen", "elmo")

people_words<-c("boy", "mom", "dad", "sister", "brother", "grandmother", "grandfather", "aunt", "uncle", "child", "teacher", "girl", "great_aunt", "zookeeper_wife", "cousin", "kid", "man")

clothes_words<-c("belt", "shorts", "shirt", "hat", "glasses", "sunglasses", "sock", "slipper", "boot", "pants", "watch", "gown", "sweater", "dress", "necklace", "pocket", "hood", "sneaker", "shoe", "coat")

nature_words<-c("tree", "flower", "windmill", "grass", "wing", "butterfly", "thunder", "lightning", "watering_can", "iris", "stick", "plant", "worm", "snail", "pond")
```

```{r}
runs <- coded_responses %>%
  select(subj, age, referent, time) %>%
  arrange(subj, age, referent, time) %>%
  group_by(subj, age, referent) %>%
  mutate(diff = time-lag(time)) %>%
  filter(diff < 10 & diff > 1)

fixed_runs <- map(1:100, function(row) {
  times <- 1:(runs[row,]$diff-1) + runs[row,]$time
  
  data_frame(time = times, subj = runs[row,]$subj, age = runs[row,]$age, 
             referent = runs[row,]$referent)}) %>%
  bind_rows() %>%
  bind_rows(runs)
```

```{r}
gleit<-fixed_runs %>%
  mutate(topic = ifelse(referent %in% food_words,  "Food", ifelse(referent %in% animal_words, "Animal", ifelse(referent %in% body_words, "Body", ifelse(referent %in% basic_words, "Basic", ifelse(referent %in% baby_words, "Baby", ifelse(referent %in% home_words, "Home", ifelse(referent %in% toy_words, "Toy", ifelse(referent %in% people_words, "Person", ifelse(referent %in% clothes_words, "Clothes", ifelse(referent %in% nature_words, "Nature", "Other"))))))))))) %>%
  filter(subj==103, age==14)#arbitrary choice

gleit_order <- gleit %>%
  ungroup() %>%
  group_by(referent) %>%
  arrange(time) %>%
  slice(1) %>%
  arrange(time) 

gleit_ordered <- gleit %>%
  mutate(referent = factor(referent, levels = gleit_order$referent)) 

ggplot(gleit_ordered, aes(x = as.factor(time), y = referent, fill=topic)) + 
  geom_tile(color = "white", show.legend=TRUE) + 
  theme_bw()+
  scale_x_discrete(breaks= c(seq(from = 1, to = 1000, by = 10)))
```

Here, we do the same computation for topics as categories as we did for referents, where we arbitrarily decide the burstiness of a topic (if it is mentioned after n times) and then fill in the gaps to see how topics shift. Some topics here seem pretty bursty (ex. food, animal), and others do not seem bursty (ex. person). Overall, kids seem to topic shift a lot! 

```{r calculating topic switches} 
##BROKEN!
topic_runs <- coded_responses %>%
  mutate(topic = ifelse(referent %in% food_words,  "Food", ifelse(referent %in% animal_words, "Animal", ifelse(referent %in% body_words, "Body", ifelse(referent %in% basic_words, "Basic", ifelse(referent %in% baby_words, "Baby", ifelse(referent %in% home_words, "Home", ifelse(referent %in% toy_words, "Toy", ifelse(referent %in% people_words, "Person", ifelse(referent %in% clothes_words, "Clothes", ifelse(referent %in% nature_words, "Nature", "Other"))))))))))) %>%
  select(subj, age, referent, time, topic) %>% #person data not included here, include?
  group_by(subj, age, referent) %>%
  arrange(subj, age, time) %>%
  group_by(subj, age, topic) %>%
  mutate(diff_topic = time-lag(time)) %>%
  filter(diff_topic < 10 & diff_topic > 1)

fixed_topic_runs <- map(1:100, function(row) {
  times <- 1:(topic_runs[row,]$diff_topic-1) + topic_runs[row,]$time
  
  data_frame(time = times, subj = topic_runs[row,]$subj, age = topic_runs[row,]$age, 
             referent = topic_runs[row,]$referent, topic = topic_runs[row,]$topic)}) %>%
  bind_rows() %>%
  bind_rows(topic_runs)

topic_gleit_order <- topic_runs %>%
  filter(subj==33, age==14) %>%
  ungroup() %>%
  group_by(topic) %>%
  arrange(time) %>%
  slice(1) %>%
  arrange(time) 

topic_gleit_ordered <- topic_runs %>%
  mutate(topic = factor(topic, levels = topic_gleit_order$topic))

#topics are all over the place!  
ggplot(topic_gleit_ordered, aes(x = as.factor(time), y = topic)) + 
  geom_tile(color = "white", show.legend=TRUE) + 
  theme_bw()+
  scale_x_discrete(breaks= c(seq(from = 1, to = 1000, by = 10)))
```

Provides general information about frequency of different topics by age. 
```{r topic trends}
topic2<- fixed_topic_runs %>%
  filter(subj==103| subj==33)

ggplot(topic2, aes(x=topic, label=topic, fill=topic))+ #colors correspond to bottom labels 
  geom_bar(position="dodge")+
  facet_grid(subj~age)+
  theme(legend.position="bottom")
```

