---
title: "Concreteness"
author: "Maddie Meyers, Dan Yurovsky"
date: "10/26/2017"
output:
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(langcog)
library(stringr) 
library(lme4)
library(directlabels)
library(DT)
library(broom)
library(data.table)
library(ggplot2)
library(forcats)
library(feather)
library(tidytext)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = TRUE, tidy = FALSE)

theme_dy <- function(base_size = 14) 
{
  theme_bw() +   
    ggplot2::`%+replace%`(ggplot2::theme_bw(base_size = base_size),
                          ggplot2::theme(panel.grid = ggplot2::element_blank(), 
                                         legend.position = "none"))
}

theme_set(theme_dy())
```

## Concreteness Judgments
We want to get a measure for how "concrete" our referents are, to protect against running into problems for words that are spoken about and not present, etc. Conreteness may also be a covariate and related to frequency in our model. 
```{r}
# Original data
loaded_subjs <- read_feather('feathers/loaded_subjs_full.feather')
# Referents and their frequencies
referents <- read_feather("feathers/referents.feather")
# End result of splitting referents and coding modality
coded_responses <- read_feather("feathers/coded_responses.feather")

#Concreteness ratings
concrete<-read_tsv("Concreteness_ratings_Brysbaert_et_al_BRM.txt") %>%
  filter(Bigram==0) %>% #take out two-word phrases
  rename(referent = Word)
```


```{r concrete}
concrete_match <-referents %>%
  mutate(match = if_else(referent %in% concrete$referent,"Match", "No Match"))

sum(concrete_match$match=="Match")
#This concreteness measure is going to be very helpful--it contains 1,266 referents from the data set and only 325 are not included
#the referents not included seem to be obscure things (taquito, spongebob, digger_truck, rice_krispie, grilled_cheese, cheeto, etc.)
```

Now we know that the concreteness data encompasses many of the referents we use. Is it a covariate for referent frequency?
```{r}
concrete_data <- referents %>%
  inner_join(concrete, by="referent")

ggplot(concrete_data, aes(x=log(freq), y=Conc.M))+
  geom_point() +
  geom_smooth()
#there don't seem to be any real trends here, try filtering a little

ggplot(filter(concrete_data, freq<500), aes(x=log(freq), y=Conc.M))+
  geom_point()+
  geom_smooth()
```

What are these things that are very low concreteness in the data?
```{r}
#there are only 56 referents here with concreteness under 4, so we do not have to really worry. A lot of these are words like angel, ghoul, etc. 
low_concrete<-concrete_data %>% 
  filter(Conc.M < 4)
```

Is concreteness related to gesture? Are people more likely to gesture to things that are more concrete?
```{r}
modality_concrete<-coded_responses %>%
  inner_join(concrete, by="referent") %>%
  mutate(modality = factor(modality, 
                           levels = c("speech", "gesture", "both")))

ggplot(modality_concrete%>%filter(Conc.M>=4), aes(x=Conc.M, fill=modality)) +
  geom_bar() #not any super obvious trends here

#Let's compare mean concreteness ratings by modality
mean_concrete<-modality_concrete %>%
  group_by(modality) %>%
  summarise(conc_modality= mean(Conc.M)) 

```

```{r}
#simple model
fit <- lm(Conc.M ~ modality, data=modality_concrete)
summary(fit)

#not simple model
fit <- glmer(cbind(modality == "speech", 
                 modality == "gesture") 
             ~ log(freq) * Conc.M + log(freq) * scale(age) +
              + (1|referent) + (1|subj), 
           data = filter(modality_concrete, person == "child"),
           family = "binomial")
```

```{r}
plot_data <- modality_concrete %>%
  filter(person == "child") %>%
  group_by(referent) %>%
  summarise(freq = mean(log(freq)),
            conc = mean(Conc.M),
            speech_prob = mean(modality == "speech"))

ggplot(plot_data, aes(x = scale(freq), y = exp(conc), color = speech_prob)) + 
  geom_point() +
  theme(legend.position = "right")
  

plot_data_2 <- modality_concrete %>%
  filter(person == "child") %>%
  #group_by(referent) %>%
  mutate(conc_bin=cut(Conc.M, 6)) %>%
  group_by(conc_bin, freq) %>%
  summarise(speech_prob = mean(modality == "speech")) 


ggplot(plot_data_2, aes(x = log(freq), y = speech_prob)) +
  facet_wrap(~ conc_bin) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic()

plot_data_3<-modality_concrete %>%
  filter(person == "child") %>%
  #group_by(referent) %>%
  group_by(freq_cut, Conc.M) %>%
  summarise(speech_prob = mean(modality == "speech"))

ggplot(plot_data_3, aes(x = Conc.M, y = speech_prob)) +
  facet_wrap(~ freq_cut) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic()

```


```{r}
ggplot(modality_concrete%>%filter(Conc.M>=4), aes(x=Conc.M))+
  geom_bar()+
  geom_vline(aes(xintercept=4.791082),  color="blue")+
  annotate(geom = "text", x = 4.830082, y = 7000, 
           label = "Both", color="blue")+
  geom_vline(aes(xintercept=4.788232),  color="green")+
  annotate(geom = "text", x = 4.7450082, y = 6000, 
           label = "Gesture", color="green")+
  geom_vline(aes(xintercept=4.765145),  color="red")+
  annotate(geom = "text", x = 4.7082, y = 5000, 
           label = "Speech", color="red")
```



