---
title: "extra_gesture_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
This is all of your work in progress code, code that failed/doesn't seem useful anymore but you don't want to delete, and questions that you have left to work on 

Questions to work on:
how do we capture a discourse? 

Right now:
1.Figure out how to split multiple referents when taking the sample data
2.Fix the issue with multi bootstrapping in the plot of everything  
3."mom" data--get at how long these repeated instances are, what is happening between them (is parent on a different referent and child is trying to get attention? How understand or define an instance of attention-getting), see if mom proportion difference by age is significant 
4.Update dan, push code through github  

```{r}
#Try making Frank et al. style Gleitman plots. It turns out we have too many different referents... Here are a few ways of collapsing the data to get a better sense for it

gleitman_data <- coded_responses %>%
  #filter(subj == 29, age == 22) %>%
  filter(subj==29) %>%
  mutate(order = as.numeric(as.factor(time))) %>%
  filter(order > 100, order < 300, freq>200) 
  

ref_freqs <- gleitman_data %>%
  group_by(referent) %>%
  summarise(n = n()) %>%
  arrange(n)

gleitman_sorted <- gleitman_data %>%
  mutate(referent = factor(referent, levels = ref_freqs$referent))
  
#plot of all the referent mentions with parent and child different modalities, sorted by frequency of referent for a specific subset  
ggplot(aes(x = order, y = referent, fill = interaction(person, modality)),
           data = gleitman_sorted) + 
  geom_tile(color = "black") +
  theme(legend.position = "bottom") +facet_wrap (~age)

discourse_data <- coded_responses %>%
  mutate(order = as.numeric(as.factor(time))) %>%
  filter(referent != "mom") %>%
  filter(subj==29, age==22) %>%
  filter(order > 25, order < 95) %>%
  mutate(new_discourse = lag(referent) == referent)%>%
  group_by(person,new_discourse, order, modality) %>%
  summarize(count=n()) %>%
  mutate(proportion=count/sum(count)) 

#plot showing tradeoff of discourse (need to choose individual subject, too much to look at overall)
ggplot(aes(x = order, y = person, fill = new_discourse), data = discourse_data) + 
  geom_tile(color = "black") +
  theme(legend.position = "bottom")+facet_wrap( ~ modality)
```

```{r modality/person transitions}
#make probability matrix of transitions between parent speech, child speech, parent gesture, child gesture, parent both, child both (AKA transition matrix of interaction between person and modality) 

transition_responses<-coded_responses %>%
  select(subj, age, person, time, referent, modality) %>%
  filter(modality != "both") %>%
  group_by(subj, age) %>%
  mutate(same_person = person == lag(person), 
         same_modality = modality == lag(modality), 
         same_referent = referent == lag(referent)) %>%
  group_by(age, person, modality, subj) %>%
  summarise_at(vars(same_person, same_modality, same_referent), funs(mean), na.rm = T) %>%
  gather(measure, value, same_person, same_modality, same_referent) %>%
  group_by(age, person, modality, measure) %>%
  summarise_at(vars(value), funs(mean, sem))


ggplot(transition_responses, aes(x = age, y = mean, color = person,
                                 label = person)) + 
  facet_grid(measure ~ modality) + 
  geom_pointrange(aes(ymin = mean - sem, ymax = mean + sem), 
                  position = position_dodge(.5)) + 
  geom_line() + 
  geom_dl(method = list(dl.trans(x=x +.2), "last.qp", cex=1)) +
  scale_x_continuous(limits = c(13, 30))

same_ref_transitions<-coded_responses %>%
  select(subj, age, person, time, referent, modality) %>%
  group_by(subj, age) %>%
  mutate(same_person = person == lag(person),
         mom_person = lag(person=="parent"),
         child_person = lag(person=="child"), 
         same_modality = modality == lag(modality), 
         same_referent = referent == lag(referent)) %>%
  filter(same_referent) %>%
  group_by(age, person, modality, subj) %>%
  summarise_at(vars(same_modality, child_person, same_person), funs(mean), na.rm = T) %>%
  gather(measure, value,  same_modality, child_person, same_person) %>%
  group_by(age, person, modality, measure) %>%
  summarise_at(vars(value), funs(mean, sem))
  
ggplot(same_ref_transitions, aes(x = age, y = mean, color = person,
                                 label = person)) + 
  facet_grid(measure ~ modality) +
  geom_pointrange(aes(ymin = mean - sem, ymax = mean + sem), 
                  position = position_dodge(.5)) + 
  geom_line() + 
  geom_dl(method = list(dl.trans(x=x +.2), "last.qp", cex=1)) +
  scale_x_continuous(limits = c(13, 30))
```

```{r}
#break up by what happens after child gestures or child speaks, who mentions next? 
speech_first <- coded_responses %>%
  group_by (subj, age, person, referent) %>%
  mutate(person_lag = lag(person)) %>%
  filter(modality)
  select(subj, age, person, modality, time, referent) %>%
  unite(type, person, modality, remove = FALSE) %>%
  mutate(child_lag=lag(type)) %>%
  group_by(subj, age, person, modality) %>%
  filter(child_lag=="child_gesture") %>%
  group_by(subj, age, person, modality) %>%
  summarise(n = n()) %>%
  mutate(n = n/sum(n)) %>% #probability
  group_by(age, person, modality) %>%
  multi_boot_standard("n")

ggplot(child_lag, aes(x = age, y = mean, color = modality)) + 
  facet_wrap(~ person) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  size=.75, position = position_dodge(1),
                  show.legend = FALSE)+
  geom_line() +   
  theme(legend.position= "bottom") + 
  ylab("proportion of utterances") 
```

```{r mom code}
mom_data<-coded_responses %>%
  group_by(subj, age, modality) %>%
  mutate(repeated= lag(referent),
         future=lead(modality)) %>%
  group_by(subj, age) %>%
  filter(referent=="mom") %>%
  filter(repeated=="mom"|repeated=="cat"|repeated=="dog"|repeated=="ball" | repeated=="baby" | repeated=="child") 


ggplot(aes(x=future, fill=modality), data=mom_data) + theme(legend.position="bottom")+geom_bar()+facet_wrap(~person) #shows that there is a lot of child repetition of the word mom


#what we have here is a data set of all the times mom was referenced, with a variable saying what the referent was before this. We can use the time column to tell us if/how many times the referent "mom" was repeated by seeing how large the time steps are between references of mom 
#granted, this is somewhat flawed because there may be other references between mom--if they are trying to get attention, the mom may say a different referent in between, so let's try going two lags back
mom_repeated <-coded_responses %>%
  mutate(time= rownames(.)) %>%
  mutate(last_ref=lag(referent),
         last_last_ref=lag(last_ref)) %>%
  filter(referent=="mom")
```

Does parent/child production of reference in *both* modalities simultaneously changes with age and word frequency?

```{r both_probabilities}
both_data <- modality_data %>%
  group_by(person, age, freq_cut) %>%
  summarise(both = sum(both), gesture = sum(gesture), speech = sum(speech)) %>%
  gather(modality, count, both, gesture, speech) %>%
  group_by(person, freq_cut, age) %>%
  mutate(count = count/sum(count)) %>%
  group_by(person, freq_cut, modality) %>%
  summarise(count = mean(count))

ggplot(aes(x = freq_cut, y = count, color = modality, group = modality), 
       data = both_data) + 
  facet_wrap(~ person) + 
  geom_point() + 
  theme(legend.position = "right") + 
  geom_smooth(se = FALSE)
```

Let's also check whether production of *both* modalities simultaneously changes over development and frequency

```{r both_model}

child_both_lm <- glmer(cbind(both, speech + gesture) ~ age * log(freq) + 
                    (1|referent) + (1|subj), 
      data = filter(modality_data, person == "child"),
      family = "binomial")

datatable(tidy_table(child_both_lm), rownames = FALSE, 
          options = list(searching = FALSE), width = 400)

parent_both_lm <- glmer(cbind(both, speech + gesture) ~ age + log(freq) + 
                    (1|referent) + (1|subj), 
      data = filter(modality_data, person == "parent"),
      family = "binomial")

datatable(tidy_table(parent_both_lm), rownames = FALSE, 
          options = list(searching = FALSE), width = 400)

```

Mom as a unique referent
```{r "mom"}
#what we have here is a data set of all the times mom was referenced, with a variable saying what the referent was before this. We can use the time column to tell us if/how many times the referent "mom" was repeated by seeing how large the time steps are between references of mom 
#granted, this is somewhat flawed because there may be other references between mom--if they are trying to get attention, the mom may say a different referent in between, so let's try going two lags back
mom_repeated <-coded_responses %>%
  mutate(last_ref=lag(referent),
         last_utt=lag(interaction(person,modality)),
         next_per=lead(person), 
         last_last_ref=lag(last_ref),
         ref=ifelse(referent=="mom", 1, 0),
         fut_utt=lead(interaction(person, modality))) %>%
  filter(person=="child") 
#make proportions
#really only interested in how the child uses "mom", not as much how the mother refers to herself
  
mom_prop<-coded_responses %>%
  filter(person=="child") %>%
  group_by(subj, age) %>%
  summarise(prop= sum(as.integer(referent=="mom"))/sum(n())) 

#when just looking at the plot of mom references across age for child, you see that there is a very sharp increase from 18 to 22 months--is this driving some of our other results? Is this difference significant? Ran a t-test and it was not significant (p=0.0885)
ggplot(aes(x=age, y=(prop/10)), data=mom_prop)+geom_bar(stat="identity")+ylab("Proportion of mom referents by child")

#how do we prove that the word "mom" is being used to get attention, versus being more communicative? look at the modalities of what precedes it
#child more likely than parent to say next referent if "mom" was previous referent, but also overall (just shows that children repeat themselves?)
#need to see what proportion are child next for mom versus for any other referent
ggplot(aes(x=age, fill=next_per), data=mom_repeated) +geom_bar()+facet_wrap(~ref)+theme(legend.position="bottom")

ggplot(aes(x=last_utt), data=mom_repeated) +geom_bar()+facet_wrap(~ref)

```

```{r first reference length}
#previous studies of parent data have shown that the first time a referent is introduced, this introduction is longer. Is this true for child-directed speech? Yes! The chats have 5 more characters
first_parent_data<-coded_responses %>%
  mutate(isFirst = ifelse(time2 %in% first_data$time2, "first","not first")) %>%
  filter(person=="parent") %>%
  mutate(utt_length=nchar(chat)) %>%
  group_by(isFirst) %>%
  summarise(sum= sum(utt_length, na.rm = TRUE) / length(isFirst))

```

