---
title: Speech/gesture tradeoffs in natural interactions
author: Dan Yurovsky, Maddie Meyers, Nicole Burke, Susan Goldin-Meadow, and Amanda Woodward
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: hide
---

```{r setup, include = FALSE}
# load packages
library(knitr)
library(tidyverse)
library(langcog)
library(stringr) 
library(lme4)
library(directlabels)
library(DT)
library(broom)
library(data.table)
library(ggplot2)
library(forcats)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = TRUE, tidy = FALSE)

theme_dy <- function(base_size = 14) 
{
  theme_bw() +   
  ggplot2::`%+replace%`(ggplot2::theme_bw(base_size = base_size),
                          ggplot2::theme(panel.grid = ggplot2::element_blank(), 
        legend.position = "none"))
}

theme_set(theme_dy())
```

## R Markdown

Load in coded data and show how many children we have at each age
```{r load_data}
files <-list.files("ldp_data/fixed_data/", 
                   "*.csv", full.names = TRUE)

loaded_subjs <- map(files, fread, sep=",", na.strings = c("NA", "")) %>%
  bind_rows() %>%
  mutate(age = 10+session*4) %>%
  group_by(subj, age) %>%
  mutate(time = 1:n()) %>%
  select(time, subj, age, person, spoken_obj, gestured_obj, referent, gloss, chat) 

demos <- loaded_subjs %>%
  select(subj,age) %>%
  distinct(subj,age, .keep_all = TRUE) %>%
  group_by(age) %>%
  summarise(n = n())

  
datatable(demos, rownames = FALSE, options = list(searching = FALSE), width = 300)
```

Let's look at the referents first

```{r referent_freqs}
# Compute the frequency of each distinct referent
referents <- loaded_subjs %>%
  ungroup() %>%
  select(referent) %>%
  filter(!is.na(referent)) %>%
  map(~str_split(string = ., pattern = ";")) %>% #split multiple-referents
  unlist() %>%
  map(str_trim) %>% # trim white space
  unlist() %>%
  data_frame(referent = .) %>%
  group_by(referent) %>%
  summarise(freq = n()) %>%
  arrange(desc(freq)) %>%
  mutate(rank = 1:nrow(.)) %>%
  select(rank,referent, freq)

referent_names <- datatable(referents, rownames = FALSE, width = 600)
referent_names

#sort referents alphabetically to try to find mistakes in data 
abc_referents <- referents[order(referents$referent),]
datatable(abc_referents, caption="Alphabetical List of Referents", 
          rownames = FALSE, width = 600)

```

Plot referent frequences in normal and log-log scale
```{r ref_plots, fig.width = 5, fig.height = 2.5}
ggplot(referents, aes(x = rank, y = freq, group = 1)) + 
  geom_point(size = .5) +
  xlab("Rank Frequency") +
  ylab("Frequency")


ggplot(referents, aes(x = rank, y = freq, group = 1)) + 
  geom_point(size = .5, color = "darkred") +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  scale_y_log10(name = "Log Frequency") +
  scale_x_log10(name = "Log Rank Frequency") + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

```{r childes_comparison, include=FALSE, eval= FALSE}
childes_freqs <- fread("ldp_data/coded_data/parentfreq.txt", sep = " ") %>%
  rename(freq = V1, word = V2) %>%
  mutate(rank = 1:nrow(.))


ggplot(childes_freqs, aes(x = rank, y = freq, group = 1)) + 
  geom_point(size = .5, color = "darkred") +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  scale_y_log10(name = "Log Frequency") +
  scale_x_log10(name = "Log Rank Frequency") + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

It looks like our referent frequency distribution is not actually well described by Zipf's law (yet?). There aren't enough low frequency referents in the tail. Let's try with just words:

Plot referent frequences in normal and log-log scale

```{r spoken_ref_plots, fig.width = 5, fig.height = 2.5}
ggplot(spoken_referents, aes(x = rank, y = freq, group = 1)) + 
  geom_point(size = .5) +
  xlab("Rank Frequency") +
  ylab("Frequency") +
  theme()


ggplot(spoken_referents, aes(x = rank, y = freq, group = 1)) + 
  geom_point(size = .5, color = "darkred") +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  scale_y_log10(name = "Log Frequency") +
  scale_x_log10(name = "Log Rank Frequency") + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

Looks similar... Maybe things will look a bit better with more data coded? Interesting to think about why this isn't Zipfian. Maybe we should think about discourse topics?

Ok, let's try the main analysis: Does frequency predict speech/gesture?

```{r speech_gesture_freq}
max_refs <- loaded_subjs %>%
  ungroup() %>%
  gather(ref_col, refs, spoken_obj, gestured_obj, referent) %>%
  mutate(num_refs = str_count(refs,";")) %>%
  summarise(num_refs = max(num_refs, na.rm = T)+1) %>%
  as.integer()
  
ref_list <- sapply(1:max_refs,function(x) paste0("referent",x))

spread_referents <- loaded_subjs %>%
  separate_("referent", ref_list, sep = ";", fill = "right") %>%
  gather_("number", "referent", ref_list) %>%
  filter(!is.na(referent)) %>%
  mutate(referent = str_trim(referent))

coded_responses <- spread_referents %>%
  mutate(spoken_obj = map(str_split(spoken_obj, ";"), str_trim),
         gestured_obj = map(str_split(gestured_obj, ";"), str_trim)) %>%
  rowwise() %>%
  mutate(spoken = purrr::contains(spoken_obj[[1]], referent),
         gestured = purrr::contains(gestured_obj[[1]], referent),
         both = spoken & gestured) %>%
  ungroup() %>%
  mutate(modality = if_else(spoken & gestured, "both", 
                            if_else(spoken, "speech", "gesture"))) %>%
  select(time, subj, age, person, referent, modality, chat, gloss) %>%
  left_join(referents) %>%
  mutate(freq_cut = cut(log(freq), 4))

modality_data <- coded_responses %>%
  group_by(person, age, subj, freq_cut, rank, freq, referent, modality, chat) %>%
  summarise(n = n()) %>%
  spread(modality, n, fill = 0)
```

Does parent/child production of reference in *both* modalities simultaneously changes with age and word frequency?

```{r both_probabilities}
both_data <- modality_data %>%
  group_by(person, age, freq_cut) %>%
  summarise(both = sum(both), gesture = sum(gesture), speech = sum(speech)) %>%
  gather(modality, count, both, gesture, speech) %>%
  group_by(person, freq_cut, age) %>%
  mutate(count = count/sum(count)) %>%
  group_by(person, freq_cut, modality) %>%
  summarise(count = mean(count))

ggplot(aes(x = freq_cut, y = count, color = modality, group = modality), 
       data = both_data) + 
  facet_wrap(~ person) + 
  geom_point() + 
  theme(legend.position = "right") + 
  geom_smooth(se = FALSE)
```


Try making Frank et al. style Gleitman plots. It turns out we have too many different referents... Here are a few ways of collapsing the data to get a better sense for it
```{r gleitman_plots}
gleitman_data <- coded_responses %>%
  #filter(subj == 29, age == 22) %>%
  filter(subj==29) %>%
  mutate(order = as.numeric(as.factor(time))) %>%
  filter(order > 100, order < 300, freq>200) 
  

ref_freqs <- gleitman_data %>%
  group_by(referent) %>%
  summarise(n = n()) %>%
  arrange(n)

gleitman_sorted <- gleitman_data %>%
  mutate(referent = factor(referent, levels = ref_freqs$referent))
  
#plot of all the referent mentions with parent and child different modalities, sorted by frequency of referent for a specific subset  
ggplot(aes(x = order, y = referent, fill = interaction(person, modality)),
           data = gleitman_sorted) + 
  geom_tile(color = "black") +
  theme(legend.position = "bottom") +facet_wrap (~age)


discourse_data <- coded_responses %>%
  mutate(order = as.numeric(as.factor(time))) %>%
  filter(referent != "mom") %>%
  filter(subj==29, age==22) %>%
  filter(order > 25, order < 95) %>%
  mutate(new_discourse = lag(referent) == referent)%>%
  group_by(person,new_discourse, order, modality) %>%
  summarize(count=n()) %>%
  mutate(proportion=count/sum(count)) 

#plot showing tradeoff of discourse (need to choose individual subject, too much to look at overall)
ggplot(aes(x = order, y = person, fill = new_discourse), data = discourse_data) + 
  geom_tile(color = "black") +
  theme(legend.position = "bottom")+facet_wrap( ~ modality)


```


```{r ref_density}

# Takes a data frame (grouped), a consecutive sample size (e.g. 10 rows), and a number of samples (e.g. 20 samples) and returns a sub-sampled dataframe with a new column "sample" that indicates the sample number

split_df <- function(df, size, count) {
 nrows <- df %>% 
   summarise(n = n()) %>% 
   ungroup() %>% 
   select(n) %>% 
   min
  
 num_splits <- floor(nrows / size)
 
 split_list <- map(1:num_splits, function(n) (n-1)*size + 1:size) %>%
   sample(., count)
 
 
 samples <- map(split_list, function(n) df %>% slice(n))
 
 sampled_df <- map(1:length(samples), 
                   function(n) mutate(samples[[n]], sample = n)) %>%
   bind_rows()
}

# try e.g. 
tmp <- coded_responses %>%
  group_by(subj, age) 

sampled_df <- split_df(tmp, 16, 20)

#only we want all of the rows in the original data, not coded responses


```

```{r "first" referent}
#here, we want to see what modality the parent or child is most likely to use to mention a new referent where new=different from the last one 

discourse_child_data <- coded_responses %>%
  mutate(order = as.numeric(as.factor(time))) %>%
  filter(referent != "mom") %>%
  mutate(new_discourse = lag(referent) == referent)%>%
  filter(person=="child") %>%
  group_by(age, new_discourse, modality) %>%
  summarize(count=n()) %>%
  mutate(proportion=count/sum(count)) 

discourse_parent_data <- coded_responses %>%
  mutate(order = as.numeric(as.factor(time))) %>%
  filter(referent != "mom") %>%
  mutate(new_discourse = lag(referent) == referent)%>%
  filter(person=="parent") %>%
  group_by(age, new_discourse, modality) %>%
  summarize(count=n()) %>%
  mutate(proportion=count/sum(count)) 

#plot of porportion based on modality and whether it was a new referent or not with age
#child
ggplot(aes(x = modality, y=proportion, fill = new_discourse),
           data = discourse_child_data ) + 
  geom_bar(position="dodge", stat="identity")+ 
  theme(legend.position = "bottom") + facet_wrap (~ age)  

#parent (what is the weird NA?)
ggplot(aes(x = modality, y=proportion, fill = new_discourse),
           data = discourse_parent_data ) + 
  geom_bar(position="dodge", stat="identity")+ 
  theme(legend.position = "bottom") + facet_wrap (~ age) 

#plot of everything on one (harder to see)
ggplot(discourse_child_data, aes(x = modality, y = proportion, color=new_discourse)) + 
  facet_wrap(~ age) + 
  geom_point()+geom_point(data=discourse_parent_data, size=3)+theme(legend.position="bottom")
```

```{r first referent}
#we want to see the probability of modality/person referencing something for the first time, not just if it is the same or different from the last referent 

first_data <- coded_responses %>%
  group_by(subj, age, referent) %>%
  slice(1) %>% 
  mutate(first = TRUE) #take first time it is referenced within the session

#take out first 
not_first_data <- coded_responses %>%
  group_by(subj, age, referent) %>%
  slice(2:nrow(.)) %>% 
  mutate(first = FALSE)

usage_data <- bind_rows(first_data, not_first_data) %>%
  filter(freq>1) %>% #taking out referents that are only mentioned once changes nothing
  group_by(subj, age, person, modality, first) %>%
  summarise(n = n()) %>%
  group_by(age, subj, person, first) %>%
  mutate(n = n/sum(n)) %>% #probability
  group_by(age, person, modality, first) %>%
  multi_boot_standard("n")

#from this plot, we can see that for the children < 26mos, they are more likely to gesture to an object if it is the first utterance versus not first utterance, and they are also more likely to both gesutre/speech to the object if it is the first utternace. They are less likely to use speech as the first reference. The confidence intervals show that these relationships are significant
ggplot(usage_data, aes(x = age, y = mean, color = modality,
                       linetype = first)) + 
  facet_wrap(~ person) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  size=.75, position = position_dodge(1),
                  show.legend = FALSE)+
  geom_line() +   
  theme(legend.position= "bottom") + 
  ylab("proportion of utterances") 


#now we are looking at frequency and seeing what happens for first or not first referents across the frequency for what modality and what person, see that parents over time increase speech (especially for first referents?)
usage_data_2 <- bind_rows(first_data, not_first_data) %>%
  filter(freq>1) %>% #taking out referents that are only mentioned once changes nothing
  group_by(subj, freq_cut, person, modality, first) %>%
  summarise(n = n()) %>%
  group_by(freq_cut, subj, person, first) %>%
  mutate(n = n/sum(n)) %>% #probability
  group_by(freq_cut, person, modality, first) %>%
  multi_boot_standard("n")

ggplot(usage_data_2,
       aes(x = as.numeric(freq_cut), y = mean, 
           color = person, linetype = first)) + 
  facet_wrap(~ modality) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  size=.75, position = position_dodge(1),
                  show.legend = FALSE)+
  geom_line(position = position_dodge(1)) +   
  theme(legend.position= "bottom") + 
  ylab("proportion of utterances") 

#break up by what happens after child gestures or child speaks, who mentions next
speech_first <- coded_responses %>%
  group_by (subj, age, person, referent) %>%
  mutate(person_lag = lag(person)) %>%
  filter(modality)
  select(subj, age, person, modality, time, referent) %>%
  unite(type, person, modality, remove = FALSE) %>%
  mutate(child_lag=lag(type)) %>%
  group_by(subj, age, person, modality) %>%
  filter(child_lag=="child_gesture") %>%
  group_by(subj, age, person, modality) %>%
  summarise(n = n()) %>%
  mutate(n = n/sum(n)) %>% #probability
  group_by(age, person, modality) %>%
  multi_boot_standard("n")

ggplot(child_lag, aes(x = age, y = mean, color = modality)) + 
  facet_wrap(~ person) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  size=.75, position = position_dodge(1),
                  show.legend = FALSE)+
  geom_line() +   
  theme(legend.position= "bottom") + 
  ylab("proportion of utterances") 

```


```{r modality/person transitions}
#make probability matrix of transitions between parent speech, child speech, parent gesture, child gesture, parent both, child both (AKA transition matrix of interaction between person and modality) 

transition_responses<-coded_responses %>%
  select(subj, age, person, time, referent, modality) %>%
  filter(modality != "both") %>%
  group_by(subj, age) %>%
  mutate(same_person = person == lag(person), 
         same_modality = modality == lag(modality), 
         same_referent = referent == lag(referent)) %>%
  group_by(age, person, modality, subj) %>%
  #summarise_at(vars(same_person, same_modality, same_referent), funs(mean), na.rm = T) %>%
  summarise_at(vars(same_person, same_modality, same_referent), funs(mean), na.rm = T) %>%
  gather(measure, value, same_person, same_modality, same_referent) %>%
  group_by(age, person, modality, measure) %>%
  summarise_at(vars(value), funs(mean, sem))


ggplot(transition_responses, aes(x = age, y = mean, color = person,
                                 label = person)) + 
  facet_grid(measure ~ modality) + 
  geom_pointrange(aes(ymin = mean - sem, ymax = mean + sem), 
                  position = position_dodge(.5)) + 
  geom_line() + 
  geom_dl(method = list(dl.trans(x=x +.2), "last.qp", cex=1)) +
  scale_x_continuous(limits = c(13, 30))

same_ref_transitions<-coded_responses %>%
  select(subj, age, person, time, referent, modality) %>%
  group_by(subj, age) %>%
  mutate(same_person = person == lag(person),
         mom_person = lag(person=="parent"),
         child_person = lag(person=="child"), 
         same_modality = modality == lag(modality), 
         same_referent = referent == lag(referent)) %>%
  filter(same_referent) %>%
  group_by(age, person, modality, subj) %>%
  summarise_at(vars(same_modality, child_person, same_person), funs(mean), na.rm = T) %>%
  gather(measure, value,  same_modality, child_person, same_person) %>%
  group_by(age, person, modality, measure) %>%
  summarise_at(vars(value), funs(mean, sem))
  
ggplot(same_ref_transitions, aes(x = age, y = mean, color = person,
                                 label = person)) + 
  facet_grid(measure ~ modality) +
  geom_pointrange(aes(ymin = mean - sem, ymax = mean + sem), 
                  position = position_dodge(.5)) + 
  geom_line() + 
  geom_dl(method = list(dl.trans(x=x +.2), "last.qp", cex=1)) +
  scale_x_continuous(limits = c(13, 30))
```

```{r speech_gesture_plot, fig.width=9, fig.height=4}
plotting_data <- coded_responses %>%
  filter(modality != "both") %>%
  group_by(person, age, subj, freq_cut, modality) %>%
  summarise(n = n()) %>%
  mutate(prob = n/sum(n)) %>%
  group_by(person, age, modality, freq_cut) %>%
  multi_boot_standard("prob")

ggplot(plotting_data, aes(x = as.numeric(freq_cut), y = mean, color = modality,
                          label = modality)) +
  facet_grid(person ~ age) + 
  geom_pointrange(aes(ymax = ci_upper, ymin = ci_lower), 
                  position = position_dodge(.25)) + 
  scale_color_brewer(palette = "Set1") +
  geom_dl(method = list(dl.trans(x=x +.2), "last.qp", cex=1)) +
  scale_x_continuous(limits = c(.5, 5.7),
                     breaks = seq(1, 4),
                     name = "Frequency Quartile") +
  ylab("Production Probability")
```

Now lets fit some statistical models to this to see if the effects are robust

```{r fit_models}
tidy_table <- function(mermod) {
  tidy(mermod) %>%
    filter(group == "fixed") %>%
    select(-group)
}

child_lm <- glmer(cbind(speech, gesture) ~ age * log(freq) + 
                    (1|referent) + (1|subj), 
      data = filter(modality_data, person == "child"),
      family = "binomial")

datatable(tidy_table(child_lm), rownames = FALSE, 
          options = list(searching = FALSE), width = 400)

parent_lm <- glmer(cbind(speech, gesture) ~ age + log(freq) + 
                    (1|referent) + (1|subj), 
      data = filter(modality_data, person == "parent"),
      family = "binomial")

datatable(tidy_table(parent_lm), rownames = FALSE, 
          options = list(searching = FALSE), width = 400)
```

Let's also check whether production of *both* modalities simultaneously changes over development and frequency

```{r both_model}

child_both_lm <- glmer(cbind(both, speech + gesture) ~ age * log(freq) + 
                    (1|referent) + (1|subj), 
      data = filter(modality_data, person == "child"),
      family = "binomial")

datatable(tidy_table(child_both_lm), rownames = FALSE, 
          options = list(searching = FALSE), width = 400)

parent_both_lm <- glmer(cbind(both, speech + gesture) ~ age + log(freq) + 
                    (1|referent) + (1|subj), 
      data = filter(modality_data, person == "parent"),
      family = "binomial")

datatable(tidy_table(parent_both_lm), rownames = FALSE, 
          options = list(searching = FALSE), width = 400)

```
