---
title: Speech/gesture tradeoffs in natural interactions
author: Dan Yurovsky, Maddie Meyers, Nicole Burke, Susan Goldin-Meadow, and Amanda Woodward
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: hide
---

```{r setup, include = FALSE}
# load packages
library(knitr)
library(tidyverse)
library(tidyboot)
library(stringr) 
library(lme4)
library(directlabels)
library(DT)
library(broom)
library(data.table)
library(ggplot2)
library(forcats)
library(feather)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = TRUE, tidy = FALSE)

theme_dy <- function(base_size = 14) 
{
  theme_bw() +   
  ggplot2::`%+replace%`(ggplot2::theme_bw(base_size = base_size),
                          ggplot2::theme(panel.grid = ggplot2::element_blank(), 
        legend.position = "none"))
}

theme_set(theme_dy())
```

## R Markdown

Load in coded data and show how many children we have at each age
```{r load_data}
loaded_subjs<-read_feather('feathers/loaded_subjs_full.feather')

loaded_subjs <- map(files, fread, sep=",", na.strings = c("NA", "")) %>%
  map(., ~mutate_at(., "ref_predicted", as.character)) %>%
  bind_rows() %>%
  mutate(age = 10+session*4) %>%
  mutate(time2=rownames(.)) %>%
  group_by(subj, age) %>%
  mutate(time = 1:n()) %>%
  select(time, time2, subj, age, person, spoken_obj, gestured_obj, referent, ref_pres, ref_predicted, gloss, chat) 

fill_vals<-which(is.na(loaded_subjs$ref_predicted)==TRUE & is.na(loaded_subjs$referent)==FALSE)

demos <- loaded_subjs %>%
  select(subj,age) %>%
  distinct(subj,age, .keep_all = TRUE) %>%
  group_by(age) %>%
  summarise(n = n())

  
datatable(demos, rownames = FALSE, options = list(searching = FALSE), width = 300)
```

Let's look at the referents first

```{r referent_freqs}
# Compute the frequency of each distinct referent
referents <- subjs %>%
  ungroup() %>%
  select(referent) %>%
  filter(!is.na(referent)) %>%
  map(~str_split(string = ., pattern = ";")) %>% #split multiple-referents
  unlist() %>%
  map(str_trim) %>% # trim white space
  unlist() %>%
  data_frame(referent = .) %>%
  group_by(referent) %>%
  summarise(freq = n()) %>%
  arrange(desc(freq)) %>%
  mutate(rank = 1:nrow(.)) %>%
  select(rank,referent, freq)

referent_names <- datatable(referents, rownames = FALSE, width = 600)
write_feather(referents, "feathers/referents.feather")

#sort referents alphabetically to try to find mistakes in data 
abc_referents <- referents[order(referents$referent),]
datatable(abc_referents, caption="Alphabetical List of Referents", 
          rownames = FALSE, width = 600)

```

Plot referent frequences in normal and log-log scale
```{r ref_plots, fig.width = 5, fig.height = 2.5}
ggplot(referents, aes(x = rank, y = freq, group = 1)) + 
  geom_point(size = .5) +
  xlab("Rank Frequency") +
  ylab("Frequency")


ggplot(referents, aes(x = rank, y = freq, group = 1)) + 
  geom_point(size = .5, color = "darkred") +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  scale_y_log10(name = "Log Frequency") +
  scale_x_log10(name = "Log Rank Frequency") + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

```{r childes_comparison, include=FALSE, eval= FALSE}
childes_freqs <- fread("ldp_data/coded_data/parentfreq.txt", sep = " ") %>%
  rename(freq = V1, word = V2) %>%
  mutate(rank = 1:nrow(.))


ggplot(childes_freqs, aes(x = rank, y = freq, group = 1)) + 
  geom_point(size = .5, color = "darkred") +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  scale_y_log10(name = "Log Frequency") +
  scale_x_log10(name = "Log Rank Frequency") + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

It looks like our referent frequency distribution is not actually well described by Zipf's law (yet?). There aren't enough low frequency referents in the tail. Let's try with just words:

Plot referent frequences in normal and log-log scale

```{r spoken_ref_plots, fig.width = 5, fig.height = 2.5}
spoken_referents <- loaded_subjs %>%
  ungroup() %>%
  select(spoken_obj) %>%
  filter(!is.na(spoken_obj)) %>%
  map(~str_split(string = ., pattern = ";")) %>% #split multiple-referents
  unlist() %>%
  map(str_trim) %>% # trim white space
  unlist() %>%
  data_frame(referent = .) %>%
  group_by(referent) %>%
  summarise(freq = n()) %>%
  arrange(desc(freq)) %>%
  mutate(rank = 1:nrow(.)) %>%
  select(rank,referent, freq)

ggplot(spoken_referents, aes(x = rank, y = freq, group = 1)) + 
  geom_point(size = .5) +
  xlab("Rank Frequency") +
  ylab("Frequency") +
  theme()


ggplot(spoken_referents, aes(x = rank, y = freq, group = 1)) + 
  geom_point(size = .5, color = "darkred") +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  scale_y_log10(name = "Log Frequency") +
  scale_x_log10(name = "Log Rank Frequency") + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

Looks similar... Maybe things will look a bit better with more data coded? Interesting to think about why this isn't Zipfian. Maybe we should think about discourse topics?

Ok, let's try the main analysis: Does frequency predict speech/gesture?

```{r speech_gesture_freq}
max_refs <- loaded_subjs %>%
  ungroup() %>%
  gather(ref_col, refs, spoken_obj, gestured_obj, referent) %>%
  mutate(num_refs = str_count(refs,";")) %>%
  summarise(num_refs = max(num_refs, na.rm = T)+1) %>%
  as.integer()

max_ref_predicted <- loaded_subjs %>%
  ungroup() %>%
  gather(ref_predicted_col, ref_predicted, ref_predicted) %>%
  mutate(num_ref_predicted = str_count(ref_predicted,";")) %>%
  summarise(num_ref_predicted = max(num_ref_predicted, na.rm = T)+1) %>%
  as.integer()
  
ref_list <- sapply(1:max_refs,function(x) paste0("referent",x))
ref_predicted_list <- sapply(1:max_ref_predicted,function(x) paste0("ref_predicted",x))

spread_referents <- loaded_subjs %>%
  separate_("referent", ref_list, sep = ";", fill = "right") %>%
  gather_("number", "referent", ref_list) %>%
  filter(!is.na(referent)) %>%
  mutate(referent = str_trim(referent)) 

coded_responses <- spread_referents %>%
  mutate(spoken_obj = map(str_split(spoken_obj, ";"), str_trim),
         gestured_obj = map(str_split(gestured_obj, ";"), str_trim)) %>%
  rowwise() %>%
  mutate(spoken = purrr::has_element(spoken_obj[[1]], referent),
         gestured = purrr::has_element(gestured_obj[[1]], referent),
         both = spoken & gestured) %>%
  ungroup() %>%
  mutate(modality = if_else(spoken & gestured, "both", 
                            if_else(spoken, "speech", "gesture"))) %>%
  select(-number) %>%
  separate_("ref_predicted", ref_predicted_list, sep = ";", fill = "right") %>%
  gather_("number", "ref_predicted", ref_predicted_list) %>%
  filter(!is.na(ref_predicted)) %>%
  mutate(ref_predicted=str_trim(ref_predicted)) %>%
  select(time, time2, subj, age, person, referent, modality, ref_predicted, chat, gloss) %>%
  left_join(referents) %>%
  mutate(freq_cut = cut(log(freq), 4))

write_feather(coded_responses, "feathers/coded_responses.feather")

modality_data <- coded_responses %>%
  filter(ref_predicted=="1") %>%
  group_by(person, age, subj, freq_cut, rank, freq, referent, modality, chat) %>%
  summarise(n = n()) %>%
  spread(modality, n, fill = 0)
```

These plots show the tradeoff between speech and gesture across ages for children and parents. 
```{r speech_gesture_plot, fig.width=9, fig.height=4}
plotting_data <- coded_responses %>%
  filter(ref_predicted == "1") %>% #filters out referents that are not present in the environment
  group_by(person, age, subj, freq_cut, modality) %>%
  summarise(n = n()) %>%
  mutate(prob = n/sum(n)) %>%
  group_by(person, age, modality, freq_cut) %>%
  tidyboot_mean(prob)

ggplot(plotting_data, aes(x = as.numeric(freq_cut), y = mean, color = modality,
                          label = modality)) +
  geom_line()+
  facet_grid(person ~ age) + 
  geom_pointrange(aes(ymax = ci_upper, ymin = ci_lower), 
                 position = position_dodge(.25)) + 
  scale_color_brewer(palette = "Set1") +
  geom_dl(method = list(dl.trans(x=x +.2), "last.qp", cex=1)) +
  scale_x_continuous(limits = c(.5, 5.7),
                     breaks = seq(1, 4),
                     name = "Frequency Quartile") +
  ylab("Production Probability")
```

Now lets fit some statistical models to this to see if the effects are robust

```{r fit_models}
tidy_table <- function(mermod) {
  tidy(mermod) %>%
    filter(group == "fixed") %>%
    select(-group)
}

child_lm <- glmer(cbind(speech, gesture) ~ age * log(freq) + 
                    (1|referent) + (1|subj), 
      data = filter(modality_data, person == "child"),
      family = "binomial")

datatable(tidy_table(child_lm), rownames = FALSE, 
          options = list(searching = FALSE), width = 400)

parent_lm <- glmer(cbind(speech, gesture) ~ age + log(freq) + 
                    (1|referent) + (1|subj), 
      data = filter(modality_data, person == "parent"),
      family = "binomial")

datatable(tidy_table(parent_lm), rownames = FALSE, 
          options = list(searching = FALSE), width = 400)
```

Let's also check whether production of *both* modalities simultaneously changes over development and frequency

```{r both_probabilities}
both_data <- modality_data %>%
  group_by(person, age, freq_cut) %>%
  summarise(both = sum(both), gesture = sum(gesture), speech = sum(speech)) %>%
  gather(modality, count, both, gesture, speech) %>%
  group_by(person, freq_cut, age) %>%
  mutate(count = count/sum(count)) %>%
  group_by(person, freq_cut, modality) %>%
  summarise(count = mean(count))

ggplot(aes(x = freq_cut, y = count, color = modality, group = modality), 
       data = both_data) + 
  facet_wrap(~ person) + 
  geom_point() + 
  theme(legend.position = "right") + 
  geom_smooth(se = FALSE)
```

```{r both_model}
child_both_lm <- glmer(cbind(both, speech + gesture) ~ age * log(freq) + 
                    (1|referent) + (1|subj), 
      data = filter(modality_data, person == "child"),
      family = "binomial")

datatable(tidy_table(child_both_lm), rownames = FALSE, 
          options = list(searching = FALSE), width = 400)

parent_both_lm <- glmer(cbind(both, speech + gesture) ~ age + log(freq) + 
                    (1|referent) + (1|subj), 
      data = filter(modality_data, person == "parent"),
      family = "binomial")

datatable(tidy_table(parent_both_lm), rownames = FALSE, 
          options = list(searching = FALSE), width = 400)

```

## First Referent, Person, Modality Interaction

Here, we want to see how the age of the subject is related to how they introduce a new (or not new) referent into the discourse.

```{r first_modality interaction}
first_data <- coded_responses %>%
  group_by(subj, age, referent) %>%
  slice(1) %>% 
  mutate(first = TRUE) 

not_first_data <- coded_responses %>%
  group_by(subj, age, referent) %>%
  slice(2:nrow(.)) %>% 
  mutate(first = FALSE)

usage_data <- bind_rows(first_data, not_first_data) %>%
  filter(freq>1) %>% #taking out referents that are only mentioned once changes nothing
  group_by(subj, age, person, modality, first) %>%
  summarise(n = n()) %>% 
  group_by(age, subj, person, first) %>%
  mutate(n = n/sum(n)) %>% #probability
  group_by(age, person, modality, first) %>%
  tidyboot_mean(n) 

#Good Plot
ggplot(usage_data, aes(x = age, y = mean, color = modality,
                       linetype = first)) + 
  facet_wrap(~ person) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  size=.75, position = position_dodge(1),
                  show.legend = FALSE)+
  geom_line() +   
  theme(legend.position= "bottom") + 
  ylab("proportion of utterances") 

```

